Open-Set Domain Adaptation through Self-Supervision

### 2022-02-12 paper/main.tex
addded related works

%------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}

\section{Method}
\label{sec:method}

@@

Table \ref{tab:results} contains the results for all 12 available domain shifts,
as well as perfomance differences when using multi-head rotation classifiers and center loss.
Further details are to be found in section \ref{sec:implementation_details}.
We can notice that multi-headed models are more capable of correctly discriminating between known and unknown samples,
ultimately resulting in better domain alignment.
A further advantage is given when combined with center loss.
AUROC values are computed at the end of the separation stage, 
while HOS values are sampled at regular times on the entire original target dataset.

>>>>

%------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}
\textbf{Closed-Set Domain Adaptation}
Unsupervised domain adaptation is the setting where labeled training data is 
available on a source domain, but the goal is to have good performance on a 
target domain with only unlabeled data and with a different marginal distribution
of data from the source dataset. The aim is to align the learned representations 
of the source and target domains.

Methods for unsupervised domain adaptation in computer vision can be divided into 
three broad classes. The first class aims to induce alignment between the
source and the target domains in some feature space optimizing for some
measurement of distributional discrepancy. \cite{bousmalis2016domain,long2015learning,long2017deep,ganin2015unsupervised,hoffman2017cycada}

The second class of methods directly transforms the source images to resemble 
the target images with generative models \cite{taigman2016unsupervised,tzeng2017adversarial,bousmalis2017unsupervised}. 
While similar to the first class in the philosophy of alignment, 
these methods operate on image pixels directly instead of an intermediate 
representation space, and therefore can benefit from an additional
round of adaptation in some representation space. Since they rely on generative 
models, large-scale datasets must train the generative model robustly.
Such methods often require complex methods of optimization.

The third class of methods uses a model trained on the labeled source data to 
estimate labels on the target data, then trains on some of those estimated 
pseudo-labels (e.g. the most confident ones), therefore bootstrapping through the 
unlabeled target data. Sometimes called self-ensembling \cite{french2018selfensembling}, this technique is borrowed from semi-supervised learning, 
where it is called co-training \cite{saito2017asymmetric,zou2018domain,chen2017rethinking}

In contrast, the presented method uses a Self-Supervised approach, already presented in 
the recent years by many other works \cite{sun2019unsupervised,yue2021prototypical,saito2020universal,BAFFOUR2022126,kim2020crossdomain,ieeeopensetselfsup} . Self-supervised methods use pretext tasks 
(e.g., image rotation, jigsaw puzzle \cite{CarlucciJigsaw}, mutual information (MI) \cite{selfsupfeng}, 
and instance discrimination \cite{chen2020simple}) to learn high-level feature representation in 
source and target domains by jointly training a shared feature extractor.

\textbf{Open Set Recognition}
The objective of Open Set Recognition (OSR) is to learn a classifier that can 
reject the unknown samples while classifying the known classes accurately.

The problem was first formulated in \cite{firstPaperOSR}, where authors 
pointed out the possibility of an open-set sample generating a very high 
activation score for one of the known class categories. 
Since then, several other works have analyzed this challenge in the context 
of deep networks \cite{perera2019deep},\cite{hein2019relu}

In \cite{bendale2015open}, OpenMax was proposed as an alternative to the SoftMax operator using 
a statistical method apportion class probabilities to the open-set class. 
A k + 1 classifier for a k class problem was used where the extra class was treated
as the open-set class.

In \cite{ge2017generative}, a Generative Adversarial Network (GAN) based framework was used to estimate
open-set class activations.

\cite{kong2021opengan} achieve strong OSR performance by using an adversarially 
trained discriminator to delineate closed from open-set images,
leveraging real open-set images for model selection

Other approaches include reconstruction based methods 
\cite{yoshihashi2019classificationreconstruction}
which use poor test-time reconstruction as an open-set indicator, 
and prototype-based methods \cite{shu2020podn} 
which represent known classes with learned prototypes, and identify open-set images
based on distances to the prototypes.

In the recent years different works to a Self-Supervised solution, used also in the 
presented method, engaging in learning representations that are invariant to the 
transformations of the input data and improving separation of classes from each other 
and from open-set samples \cite{dhamija2021selfsupervised,jia2021selfsupervised,Perera_2020_CVPR,GeoTran}

\textbf{Self-supervised Learning}
The emerging field of self-supervised learning uses the machinery of supervised 
learning on problems where external supervision is not available. 
The idea is to use data itself as supervision for auxiliary (also called “pretext”) 
tasks that learn deep feature representations which will hopefully be informative 
for downstream “real” tasks.
Many such auxiliary tasks have been proposed in the literature, including
image colorization \cite{zhang2016colorful,larsson2017learning}, patch location 
prediction \cite{doersch2016unsuperviseddoersch2017multitask},
image jigsaw puzzle [\cite{noroozi2017unsupervised}, image inpainting \cite{pathak2016context},
contrastive learning \cite{oord2019representation}, geometric transformations \cite{dosovitskiy2015discriminative}. Another geometric 
transformation task is rotation recognition \cite{OldROS}, where input images
are rotated by multiples of 90° and the network is trained to predict the rotation
angle pf each image. This pretext task has been successfully used for both anomaly
detection \cite{golan2018deep} and closed-set domain adaptation \cite{selfsupxu}

\textbf{Open Set Domain Adaptation}
Open Set Domain Adaptation (OSDA) is a challenging domain adaptation setting 
which allows the existence of unknown classes on the target domain.
The term "OSDA" was first introduced by Busto and Gall \cite{firstosda} that
considered the setting where each domain contains, in addition to the shared
categories, a set of private categories. The currently accepted defiition of OSDA
was introduced by Saito et al. \cite{saito2018open} that considered the target as 
containing all the source categories and additional set of private categories 
that should be considered unknown, presenting Open Set
Back-Propagation (OSBP) \cite{saito2018open} as an adversarial method 
to increase the prediction variances so that the generator can choose to align 
a target sample with the source known or reject it as an unknown target samples.
Separate To Adapt (STA) \cite{sat} train a multi-binary classifier to progressively 
separate the samples of unknown and known classes.
Universal Adaptation Network (UAN) \cite{uan}, originally proposed for the universal 
domain adaptation setting that is a superset of OSDA, quantify the sample-level 
transferability and recognize the unknown samples based on the transferability.
Attract or Distract (AoD) \cite{feng2019attract} starts with a mild alignment with 
a procedure similar to \cite{saito2018open} refining the decision by using 
metric learning to reduce the intra-class distance in known classes and push 
the unknown class away from the known classes.
Jing \etal \cite{Jing_Li_Zhu_Ding_Lu_Yang_2021} have recently proposed a method
to recognize the unknown samples according to the centroid deviation angles and
employing the statistical Extreme Value Theory to recognize the unknown 
samples that are misclassified into known classes.
The approach that we are going to present is instead based on a self-supervision 
task, the rotation recognition, used to both separate known and unknown target 
samples and align the known source and target distributions.
\section{Method}
\label{sec:method}

@@

Table \ref{tab:results} contains the results for all 12 available domain shifts,
as well as perfomance differences when using multi-head rotation classifiers and center loss,
as described in the previous section.
Multi-headed models are more capable of correctly discriminating between known and unknown samples,
ultimately resulting in better domain alignment.
This effect is amplified when also using center loss.
AUROC values are computed at the end of the separation stage, 
while HOS values are sampled at regular times on the entire original target dataset.
